---
title: "A00275664_RMarkdown_3"
author: "A00275664"
date: "2023-11-30"
output: pdf_document
---

# RMarkdown Document

## Question 3

### Importing and Analysing the Data
```{r}
CarsDataset <- read.csv("C:/Users/erink/OneDrive/Desktop/Interpretation of Data/CarsDataset.csv")

names(CarsDataset)
```

```{r eval = FALSE}
str(CarsDataset) #id, price, year, cylinders, odometer, lat, long, posting_date - wrong formats
```

```{r, include= FALSE}
library(readr)
library(tidyr)
library(dplyr)
library(stringr)
library(lubridate)
```

## Handling Null Values

```{r}
CarsDataset1a<-CarsDataset
#Replaces Blank Spaces with NA, if any
CarsDataset1a <- mutate(CarsDataset1a, 
                        id= na_if(id, ""),
                        url= na_if(url, ""),
                        region= na_if(region, ""),
                        region_url= na_if(region_url, ""),
                        price= na_if(price, ""),
                        year= na_if(year, ""),
                        manufacturer= na_if(manufacturer, ""),
                        model= na_if(model, ""),
                        condition= na_if(condition, ""),
                        cylinders= na_if(cylinders, ""),
                        fuel= na_if(fuel, ""),
                        odometer= na_if(odometer, ""),
                        title_status= na_if(title_status, ""),
                        transmission= na_if(transmission, ""),
                        VIN= na_if(VIN, ""),
                        drive= na_if(drive, ""),
                        size= na_if(size, ""),
                        type= na_if(type, ""),
                        paint_color= na_if(paint_color, ""),
                        image_url= na_if(image_url, ""),
                        description= na_if(description, ""),
                        state= na_if(state, ""),
                        lat= na_if(lat, ""),
                        long= na_if(long, ""),
                        posting_date= na_if(posting_date, ""))

```

Column total reduced from 42547 to 42446

## THE ID COLUMN

All the columns in the Cars Dataset have data jumbled within it that is not
relevent to its column
To remove the data, I will convert it to a character then numeric form which
will turn it into an NA value
```{r}
#Convert to numeric
Cars_Dataset1b<-CarsDataset1a
Cars_Dataset1b$id <- as.numeric(as.character(Cars_Dataset1b$id))

# Remove NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(id))

#Left with 42400 rows of data
```


## THE URL COLUMN

```{r}
# Using stringr, remove all values that do not begin with "https://
Cars_Dataset1b <- Cars_Dataset1b[str_detect(Cars_Dataset1b$url, "https://"), ]

#Changing all values to lower case
Cars_Dataset1b$url <- str_to_lower(Cars_Dataset1b$url)

# Remove all NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(url))

#Left with 42399 rows of data
```

\hfill\break
\hfill\break

## THE REGION COLUMN

```{r}
# Changing all values to lower case
Cars_Dataset1b$region <- str_to_lower(Cars_Dataset1b$region)

# Remove all NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(region))
```


## THE REGION_URL COLUMN

```{r}
# Changing all values to lower case
Cars_Dataset1b$region_url <- str_to_lower(Cars_Dataset1b$region_url)

# Using stringr, remove all values that do not begin with "https://
Cars_Dataset1b <- Cars_Dataset1b[str_detect(Cars_Dataset1b$region_url, "https://"), ]

# Remove all NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(region_url))

```

## THE PRICE COLUMN 

```{r}
# Change to numeric
Cars_Dataset1b$price <- as.numeric(as.character(Cars_Dataset1b$price))

#Remove prices that are 0
Cars_Dataset1b <- Cars_Dataset1b %>%
  filter(price != 0)

# Remove NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(price))

#Left with 39727 rows of data
```

## THE YEAR COLUMN

```{r}
# Change to numeric for better analysis
Cars_Dataset1b$year <- as.numeric(as.character(Cars_Dataset1b$year))

#Filter any year that is a 0, if any
Cars_Dataset1b <- Cars_Dataset1b %>%
  filter(year != 0)

# Remove NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(year))

```

## THE MANUFACTURER COLUMN

```{r}
# Changing all values to lower case
Cars_Dataset1b$manufacturer <- str_to_lower(Cars_Dataset1b$manufacturer)

# Replacing similiar condition variables
Cars_Dataset1b <- Cars_Dataset1b %>%
  mutate(manufacturer = case_when(manufacturer == "rover" ~ "jaguar",
                                  manufacturer == "land rover" ~ "jaguar",
                                  manufacturer == "jeep" ~ "chrysler",
                                  manufacturer == "acura" ~ "honda",
                                  manufacturer == "dodge" ~ "chrysler",
                                  manufacturer == "ram" ~ "chrysler",
                                  manufacturer == "lexus" ~ "toyota",
                                  manufacturer == "infinit" ~ "nissan",
                                  manufacturer == "mini" ~ "bmw",
                                  manufacturer == "datsun" ~ "nissan",
                                  manufacturer == "chevrolet" ~ "gmc",
                                  manufacturer == "pontiac" ~ "gmc",
                                 TRUE ~ as.character(manufacturer)))

# Remove all NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(manufacturer))
```

## THE MODEL COLUMN

```{r}
# Changing all values to lower case
Cars_Dataset1b$model <- str_to_lower(Cars_Dataset1b$model)

# Remove all NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(model))
```

## THE CONDITION COLUMN

```{r}
# Changing all values to lower case
Cars_Dataset1b$condition <- str_to_lower(Cars_Dataset1b$condition)

# Replacing similiar condition variables
Cars_Dataset1b <- Cars_Dataset1b %>%
  mutate(condition = case_when(condition == "like new" ~ "good",
                               condition == "excellent" ~ "excellent",
                               condition == "good" ~ "good",
                               condition == "fair" ~ "fair",
                               condition == "salvage" ~ "salvage",
                               condition == "new" ~ "new",
                               TRUE ~ as.character(condition)))

# Remove all NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(condition))
```

## THE CYLINDERS COLUMN 

```{r}
# Changing all values to lower case
Cars_Dataset1b$cylinders <- str_to_lower(Cars_Dataset1b$cylinders)

# Replacing cylinder variables with numeric variables
Cars_Dataset1b <- Cars_Dataset1b %>%
  mutate(cylinders = case_when(cylinders == "6 cylinders" ~ "6",
                               cylinders == "4 cylinders" ~ "4",
                               cylinders == "8 cylinders" ~ "8",
                               cylinders == "10 cylinders" ~ "10",
                               cylinders == "3 cylinders" ~ "3",
                               cylinders == "six cylinders" ~ "6",
                               cylinders == "5 cylinders" ~ "5",
                               cylinders == "12 cylinders" ~ "12",
                               cylinders == "8 cyls" ~ "8",
                               cylinders == "6 cyls" ~ "6",
                               cylinders == "4 cyls" ~ "4",
                               TRUE ~ as.character(cylinders)))


# Converting the column to a Numeric data type
Cars_Dataset1b$cylinders<- as.numeric(as.character(Cars_Dataset1b$cylinders))

# Remove all NA values created
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(cylinders))

#Left with 39631 rows of data
```


## THE FUEL COLUMN

```{r}
# Changing all values to lower case
Cars_Dataset1b$fuel <- str_to_lower(Cars_Dataset1b$fuel)

# Remove all NA values created
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(fuel))
```

## THE ODOMETER COLUMN

```{r}
# Converting the column to a Numeric data type
Cars_Dataset1b$odometer<- as.numeric(as.character(Cars_Dataset1b$odometer))

# Remove all NA values created
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(odometer))
```

## THE TITLE_STATUS COLUMN

```{r}
# Changing all values to lower case
Cars_Dataset1b$title_status <- str_to_lower(Cars_Dataset1b$title_status)

# Remove all NA values created
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(title_status))

###------INVESTIGATE UNIQUE VARIABLES-----######
```

## THE TRANSMISSION COLUMN

```{r}
# Changing all values to lower case
Cars_Dataset1b$transmission <- str_to_lower(Cars_Dataset1b$transmission)

# Remove all NA values created
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(transmission))
```

## THE VIN COLUMN #39613
There is too much data to search the VIN column for abstract data. 
I will filter out any cells that contain 0 values and null values
```{r}
Cars_Dataset1b <- Cars_Dataset1b %>%
  filter(VIN != 0)
# Remove all NA values created
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(VIN))
```

## THE DRIVE COLUMN

```{r}
# Changing all values to lower case
Cars_Dataset1b$drive <- str_to_lower(Cars_Dataset1b$drive)

# Replacing similiar variables with consistent variables
Cars_Dataset1b <- Cars_Dataset1b %>%
  mutate(drive = case_when(drive == "4wd" ~ "four wheel drive",
                           drive == "rwd" ~ "rear wheel drive",
                           drive == "fwd" ~ "front wheel drive",
                           drive == "front wheel drive" ~ "front wheel drive",
                           drive == "rear wheel drive" ~ "rear wheel drive",
                           drive == "four wheel drive" ~ "four wheel drive",
                           TRUE ~ as.character(drive)))

# Remove all NA values created
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(drive))
```

## THE SIZE COLUMN

```{r}
# Changing all values to lower case
Cars_Dataset1b$size <- str_to_lower(Cars_Dataset1b$size)

# Remove all NA values created
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(size))

```

## THE TYPE COLUMN

```{r}
# Changing all values to lower case
Cars_Dataset1b$type <- str_to_lower(Cars_Dataset1b$type)

# Remove all NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(type))

```

## THE PAINT_COLOR COLUMN

```{r}
# Changing all values to lower case
Cars_Dataset1b$paint_color <- str_to_lower(Cars_Dataset1b$paint_color)

# Remove all NA values 
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(paint_color))
```

## THE IMAGE_URL COLUMN

```{r}
# Using stringr, remove all values that do not begin with "https://
Cars_Dataset1b <- Cars_Dataset1b[str_detect(Cars_Dataset1b$image_url, "https://"), ]

# Remove all NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(image_url))

```

## THE DESCRIPTION COLUMN

```{r}
# Remove all NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(description))

```

## THE STATE COLUMN

```{r}
# Changing all values to lower case
Cars_Dataset1b$state <- str_to_lower(Cars_Dataset1b$state)

# Remove all NA values 
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(state))

```

## THE LAT COLUMN

```{r}
# Converting the column to a Numeric data type
Cars_Dataset1b$lat<- as.numeric(as.character(Cars_Dataset1b$lat))

# Remove all NA values created
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(lat))

```

## THE LONG COLUMN

```{r}
# Converting the column to a Numeric data type
Cars_Dataset1b$long<- as.numeric(as.character(Cars_Dataset1b$long))

# Remove all NA values 
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(long))
```

## THE POSTING_DATE 

```{r}
# Seperate the column into three for a more detailed analysis
Cars_Dataset1b <- separate(Cars_Dataset1b, posting_date, into = c("date", "time_zone"), sep = "T")
Cars_Dataset1b <- separate(Cars_Dataset1b, time_zone, into = c("time", "timezone"), sep = "-")
```



##THE DATE COLUMN

```{r}
# Convert the data format to date using lubridate

Cars_Dataset1b$date<- ymd(Cars_Dataset1b$date)

# Remove all NA values 
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(date))
```


## THE TIME COLUMN 

```{r}
# Convert the character format of time to time format(hms)

Cars_Dataset1b$time <- hms(Cars_Dataset1b$time)

# Remove all NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(time))

```

\hfill\break
\hfill\break


## THE TIMEZONE COLUMN

```{r}
# Convert the data format to numeric format

Cars_Dataset1b$timezone<- as.numeric(as.character(Cars_Dataset1b$timezone))

# Remove all NA values
Cars_Dataset1b<-filter(Cars_Dataset1b, !is.na(timezone))


```

### Remove Duplicate Data

```{r}
# Removing all duplicates

Cars_Dataset1b<- unique(Cars_Dataset1b) # The dataset reduced from 39613 to 39598 rows
```